---
title: "Mapping in R"
format: 
  html:
    toc: true
---

<!--
Based on:

Dr. Emily Burchfield's guide
https://www.emilyburchfield.org/courses/eds/making_maps_in_r

"Geocomputation with R" (GWR) by Lovelace, Nowosad, and Muenchow
https://r.geocompx.org/
-->

## Credits

This material is based on Dr. Emily Burchfield's [excellent guide to mapping in R](https://www.emilyburchfield.org/courses/eds/making_maps_in_r) and the book [*Geocomputation with R*](https://r.geocompx.org/) by Lovelace, Nowosad, and Muenchow.

## Setup

There is not a single mapping package for R that will suit all of your needs. We'll review several useful packages today. Start by installing any packages you don't have yet.

```{r}
# if you already have these packages, try loading them outside the suppress
# function to see if there are any key messages like update suggestions

# install.packages("spDataLarge", repos = "https://geocompr.r-universe.dev")

suppressPackageStartupMessages({
  library(tidyverse)
  library(sf)
  library(tmap)
  library(spData)
  #library(spDataLarge)
  library(leaflet)
  library(terra)
})
```

## Basic Concepts

### Vector and Raster

There are two main types of spatial data: vector and raster. Vector data includes points (a pin on a map), lines (a road), and polygons (a country shape). Raster data divides some surface into equal-sized cells and displays information contained in each cell (think satellite imagery, weather maps).

### Simple Features

> Simple features is an open standard developed and endorsed by the Open Geospatial Consortium (OGC), a not-for-profit organization whose activities we will revisit in a later chapter (in Section 8.2). Simple features is a hierarchical data model that represents a wide range of geometry types. Of 18 geometry types supported by the specification, only 7 are used in the vast majority of geographic research; these core geometry types are fully supported by the R package sf (Pebesma 2018). -GWR 

The [`sf`](https://r-spatial.github.io/sf/) package implements the simple features (or simple feature access) standards. Simple features is a ISO standard (19125-1:2004) describing how spatial properties of objects are represented in computers. A feature can be anything from a single point to a set of polygons (see @fig-sf). Every `sf` object has a geometry column that describes where the feature exists on Earth. The `sf` package is comprehensive, plays nicely with the tidyverse, and is the basis of several mapping packages. 

![Simple features supported by sf, image from [*Geocomputation with R*](https://r.geocompx.org/spatial-class).](https://r.geocompx.org/figures/sf-classes.png){#fig-sf}

Let's take a look at the `world` dataset included with the `spData` package:

```{r}
glimpse(world)
class(world)
```

You can see that `world` is a sf dataframe with non-geographic country-level attributes and a final column that contains the geography information for plotting country polygons.

## Base R

We can do a lot with base R, starting with a simple call to `plot()` which plots every variable in `world`:

```{r}
plot(world)
```

We can add layers to a base plot with the `add = TRUE` parameter.

```{r}
# create asia as a union of countries in asia
asia <- world %>%
  filter(continent == "Asia") %>%
  st_union()
```

```{r}
plot(world["pop"], reset = FALSE)
plot(asia, add = TRUE, col = "red")
```
## `ggplot()`

Some folks will plot maps exclusively with base R. They are not my people 😂 You're much more likely to find me using `ggplot2`. You can wrangle and plot a sf dataframe in familiar ways.

```{r}
world %>%
  mutate(asia = case_when(
    continent == "Asia" ~ 1,
    TRUE ~ 0
  )) %>%
  ggplot(aes(fill=asia)) +
    geom_sf() +
    theme_minimal() +
    theme(legend.position = "none") 
```
`ggplot2` also makes it easy to add different spatial features as layers. Here we start with the world and add a point for Duke.

```{r}
duke <- st_sfc(st_point(c(-78.939133, 36.001465)), crs = "EPSG:4326")
duke

# here I'm not starting with the data and piping, but I could have
ggplot() +
  geom_sf(data = world) +
  geom_sf(data = duke, color = "#003087") +
  theme_minimal() +
  theme(legend.position = "none")
```



## `sf` Package


