---
title: "Mapping in R"
format: 
  html:
    toc: true
---

<!--
Based on:

Dr. Emily Burchfield's guide
https://www.emilyburchfield.org/courses/eds/making_maps_in_r

"Geocomputation with R" (GWR) by Lovelace, Nowosad, and Muenchow
https://r.geocompx.org/

[ATTACHING CHIRPS PRECIPITATION DATA TO DHS SURVEYS](https://tech.popdata.org/dhs-research-hub/posts/2024-02-04-dhs-chirps/) by Roberts and Gunther
-->

## Credits

This material is based on Dr. Emily Burchfield's [excellent guide to mapping in R](https://www.emilyburchfield.org/courses/eds/making_maps_in_r) and the book [*Geocomputation with R*](https://r.geocompx.org/) by Lovelace, Nowosad, and Muenchow.

## Setup

There is not a single mapping package for R that will suit all of your needs. We'll review several useful packages today. Start by installing any packages you don't have yet.

```{r}
# if you already have these packages, try loading them outside the suppress
# function to see if there are any key messages like update suggestions

# install.packages("spDataLarge", repos = "https://geocompr.r-universe.dev")

suppressPackageStartupMessages({
  library(tidyverse)
  library(sf)
  library(tmap)
  library(spData)
  #library(spDataLarge)
  library(leaflet)
  library(terra)
  library(stars)
  library(chirps)
  library(usethis)
  library(ggspatial)
})
```

We need to get some data:

```{r}
#| eval: false

# run this manually once

# Burkina Faso boundary shapefile from https://www.idhsdata.org/idhs/gis.shtml
usethis::use_course(url = "https://github.com/ericpgreen/glhlth562/raw/main/materials/mapping/geo_bf2003_2010.zip", destdir = getwd())
```

## Basic Concepts

### Vector and Raster

There are two main types of spatial data: vector and raster. Vector data includes points (a pin on a map), lines (a road), and polygons (a country shape). Raster data divides some surface into equal-sized cells and displays information contained in each cell (think satellite imagery, weather maps).

### Simple Features

> Simple features is an open standard developed and endorsed by the Open Geospatial Consortium (OGC), a not-for-profit organization whose activities we will revisit in a later chapter (in Section 8.2). Simple features is a hierarchical data model that represents a wide range of geometry types. Of 18 geometry types supported by the specification, only 7 are used in the vast majority of geographic research; these core geometry types are fully supported by the R package sf (Pebesma 2018). -GWR 

The [`sf`](https://r-spatial.github.io/sf/) package implements the simple features (or simple feature access) standards. Simple features is a ISO standard (19125-1:2004) describing how spatial properties of objects are represented in computers. A feature can be anything from a single point to a set of polygons (see @fig-sf). Every `sf` object has a geometry column that describes where the feature exists on Earth. The `sf` package is comprehensive, plays nicely with the tidyverse, and is the basis of several mapping packages. 

![Simple features supported by sf, image from [*Geocomputation with R*](https://r.geocompx.org/spatial-class).](https://r.geocompx.org/figures/sf-classes.png){#fig-sf}

Let's take a look at the `world` dataset included with the `spData` package:

```{r}
glimpse(world)
class(world)
```

You can see that `world` is a sf dataframe with non-geographic country-level attributes and a final column that contains the geography information for plotting country polygons.

## Base R

We can do a lot with base R, starting with a simple call to `plot()` which plots every variable in `world`:

```{r}
plot(world)
```

We can add layers to a base plot with the `add = TRUE` parameter.

```{r}
# create asia as a union of countries in asia
asia <- world %>%
  filter(continent == "Asia") %>%
  st_union()
```

```{r}
plot(world["pop"], reset = FALSE)
plot(asia, add = TRUE, col = "red")
```
## `ggplot()`

Some folks will plot maps exclusively with base R. They are not my people 😂 You're much more likely to find me using `ggplot2`. You can wrangle and plot a sf dataframe in familiar ways.

```{r}
world %>%
  mutate(asia = case_when(
    continent == "Asia" ~ 1,
    TRUE ~ 0
  )) %>%
  ggplot(aes(fill=asia)) +
    geom_sf() +
    theme_minimal() +
    theme(legend.position = "none") 
```
`ggplot2` also makes it easy to add different spatial features as layers. Here we start with the world and add a point for Duke.

```{r}
duke <- st_sfc(st_point(c(-78.939133, 36.001465)), crs = "EPSG:4326")
duke

# here I'm not starting with the data and piping, but I could have
ggplot() +
  geom_sf(data = world) +
  geom_sf(data = duke, color = "#003087") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Working with Rasters

[`terra`](https://rspatial.github.io/terra/) and `stars` have become the go-to packages for working with raster data in R. To introduce you to raster data, we'll jump into a [tutorial](https://tech.popdata.org/dhs-research-hub/posts/2024-02-04-dhs-chirps/) on working with rainfall data from ClimateSERV. We'll use the [`chirps`](https://docs.ropensci.org/chirps/) package to get some rainfall data for Burkina Faso, using a country shapefile as a mask to tell the API which raster cells we want (the ones that fall inside Burkina Faso).

```{r}
# load shapefile of Burkina Faso boarders
bf_borders <- st_read(
  here::here("geo_bf2003_2010", "geo_bf2003_2010.shp"),
  quiet = TRUE
)

# Get CHIRPS data from 2001-2010 for Burkina Faso 
bf_precip <- get_chirps(
  vect(bf_borders),
  dates = c("2001-07-19", "2001-07-21"),
  server = "CHC"
)

bf_precip <- mask(bf_precip, bf_borders, touches = FALSE)
```

`bf_precip` is a `SpatRaster` object from the `terra` package containing the raster stack of precipitation data for Burkina Faso from three days in 2001 (using the full time series in their tutorial would take too long to download).

```{r}
bf_precip
```

> This raster layer contains 114 rows and 158 columns of pixels. The value in each pixel represents the rainfall (in millimeters) for an area 0.05 degrees longitude by 0.05 degrees latitude (shown in the resolution field).

Let's plot the rainfall on July 19, 2001. To do this we'll extract the raster for this day and plot with `ggplot()` and some functions from the `ggspatial` package (always multiple ways to do something!).

```{r}
# Create map of rainfall for single day
precip_day <- bf_precip[["chirps-v2.0.2001.07.19"]]

ggplot() +
  ggspatial::layer_spatial(
    precip_day, 
    alpha = if_else(values(precip_day) == 0, 0, 0.8), 
    na.rm = TRUE
  ) +
  ggspatial::layer_spatial(bf_borders, fill = NA, color = "#888888") +
  #ggspatial::layer_spatial(bf_gps,  fill = NA, size = 2, alpha = 0.4) +
  labs(
    title = "Burkina Faso Rainfall: July 19, 2001",
    subtitle = "CHIRPS precipitation data",
    fill = "Rainfall total (mm)",
    caption = "Source: DHS Program and Climate Hazards Center InfraRed Precipitation with Station (CHIRPS)"
  ) +
  ggspatial::annotation_scale(
    aes(style = "ticks", location = "br"), 
    text_col = "#999999",
    line_col = "#999999"
  ) +
  scale_fill_gradient(low = "white", high = "#00263A", na.value = NA) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      label.hjust = 0.5,
      barwidth = 15,
      barheight = 0.5,
      ticks.colour = "white",
      frame.colour = "#999999"
    )
  )
```

`terra` makes it easy to do raster math. Here we'll calculate the mean rainfall for these three days.

::: {.aside}
If you are following their tutorial, I commented out the line adding DHS cluster coordinates because to get this data you need to register with the DHS program.
:::

```{r}
bf_precip_mean <- mean(bf_precip)
bf_precip_mean
```

And plot

```{r}
ggplot() +
  ggspatial::layer_spatial(bf_precip_mean, alpha = 0.8, na.rm = TRUE) +
  ggspatial::layer_spatial(bf_borders, alpha = 0) +
  labs(
    title = "Burkina Faso Average Rainfall: July 19-21, 2001",
    subtitle = "CHIRPS precipitation data",
    fill = "3-day average rainfall (mm/day)",
    caption = "Source: Climate Hazards Center InfraRed Precipitation with Station (CHIRPS)"
  ) +
  ggspatial::annotation_scale(
    aes(style = "ticks", location = "br"), 
    text_col = "#999999",
    line_col = "#999999"
  ) +
  scale_fill_gradient(low = "#FAEFD1", high = "#00263A", na.value = NA) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      label.hjust = 0.5,
      barwidth = 15,
      barheight = 0.5,
      ticks.colour = "white",
      frame.colour = "#999999"
    )
  )
```
Check out the [rest of the tutorial](https://tech.popdata.org/dhs-research-hub/posts/2024-02-04-dhs-chirps/) a more depth tour.

## `sf` Package

Just as you can wrangle tibbles to filter and join records, you can also manipulate simple features with `sf` functions.

```{r}
nc <- st_read(system.file("shape/nc.shp", package="sf"))
plot(nc)
```

First, if we wanted to drop the geometry and just work with the non-geographic attributes:

```{r}
nc %>% 
  select(NWBIR74) %>%
  st_drop_geometry() 
```
We can subset geographies of interest by looking for features that intersect Ashe county (including Ashe).

```{r}
ashe <- nc %>%
  filter(NAME == "Ashe")

nc %>% 
  filter(lengths(st_intersects(., ashe)) > 0) %>%
  plot()
```

These transformation functions play nicely with tidyverse code you know and love.

```{r}
nc %>% 
  mutate(ashe_group = if_else(lengths(st_intersects(., ashe)) > 0, 
                              "Ashe+", 
                              "Rest")) %>%
  group_by(ashe_group) %>% 
  summarize(sid = sum(SID74),
            births = sum(BIR74),
            sid_births = sid/births)
```

