---
title: "#TidyTuesday"
subtitle: "TT Week 4 / Class Week 3"
author: "Eric Green"
css: custom_report.css
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 1
    code_folding: show
    theme: default
    highlight: tango
    fig_captions: no
    number_sections: false
---



```{r setup, include=FALSE, warning = FALSE, message = FALSE}
  knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
  options(scipen=999) # turns off scientific notation

# load packages
  library(tidyverse)
  library(tidytuesdayR)   # install.packages("tidytuesdayR") to get new data

# load data
  tuesdata <- tidytuesdayR::tt_load('2022-01-25')
  ratings <- tuesdata$ratings
  details <- tuesdata$details
```


# Exploration
 
*Use what you know so far to explore the dataset. If you are just starting out, write about what you want to do with the data. Try to stretch a bit every week.*

```{r}
  ratings %>%
    filter(year != 0) %>%
    group_by(year) %>%
    summarise(mean_of_avg_ratings = mean(average, na.rm=TRUE),
              games = n()) %>%
    filter(games >= 10, year >= 1958) %>%
  
    ggplot(aes(x = year, y = mean_of_avg_ratings)) +
      geom_col() +
      theme_minimal() +
      labs(title = "Recent games are rated more highly",
           subtitle = "Mean of average ratings for games introduced since 1958",
           y = NULL,
           x = NULL) +
      theme(plot.title = element_text(face = "bold"))
      
```


# Review

## Find and run

*Search the Twitter hashag [#TidyTuesday](https://twitter.com/hashtag/TidyTuesday?src=hashtag_click) for an entry that catches your eye (and includes a link to their code). Include links to the details and run their code.*

| Name | Details |
|:--   |:--      |
| Author | Replace |
| Tweet | [url](https://twitter.com/...) |
| Code | [url](https://github.com/...) |

```{r entry}
# insert their code

```

## Notes

*What did you learn from their approach?*
