---
title: "Relational Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(feb2)
```

## How many prognosticators have no prediction data?

Return a number (in a tibble is fine)

```{r}
ans <- prognosticators %>%
  anti_join(predictions, by="prognosticator_name") %>%
  count()
```

There are `r ans` prognosticators with no prediction data.


## What are their names?

Return a vector of names

```{r}
ans_names <- prognosticators %>%
  anti_join(predictions, by="prognosticator_name") %>%
  pull(prognosticator_name)

ans_names <- paste(ans_names, collapse = ", ")
```

Their names are: `r ans_names`.

## Which prognosticator's city coordinates are closest to a GHCND weather station?

Punxsutawney Phil or Jimmy the Groundhog?

```{r}
prognosticators %>%
  left_join(weather_stations_ghcnd, by = "prognosticator_city") %>%
  filter(prognosticator_name %in% c("Punxsutawney Phil", 
                                    "Jimmy the Groundhog")) %>%
  group_by(prognosticator_name) %>%
  summarize(min_distance = min(distance))
```


## Create a plot of prediction accuracy over time by `prognosticator_status`

* exclude human mascots
* limit to years with at least 5 per year per level of `prognosticator_status` (except human mascots)
* facet plot by `prognosticator_status`

```{r}
predictions %>%
# join data
  select(prognosticator_name, year, prediction) %>%
  right_join(select(filter(prognosticators, 
                           prognosticator_status!="Human Mascot"),
                    prognosticator_name, prognosticator_city,
                    prognosticator_status),
             by = "prognosticator_name") %>%
  left_join(class_def1, by = c("prognosticator_city", "year")) %>%
  filter(complete.cases(.)) %>%
# construct outcome
  mutate(result = case_when(
    prediction == class ~ "Correct",
    TRUE ~ "Incorrect"
  )) %>%
# count results by year and status
  group_by(year, prognosticator_status, result) %>%
  count() %>%
  ungroup() %>%
  complete(year, prognosticator_status, result, 
           fill = list(n=0)) %>%
# count total predictions made by year and status
  group_by(year, prognosticator_status) %>%
  mutate(n_predictions_status = sum(n)) %>%
  ungroup() %>%
# remove years if either status had <5 predictions
  group_by(year) %>%
  mutate(min = min(n_predictions_status)) %>%
  ungroup() %>%
  filter(min >= 5) %>%
# calculate percent by year, status, and result
  mutate(prop_year_status = n/n_predictions_status) %>%
# limit to correct predictions
  filter(result=="Correct") %>%
# plot
  select(year, prognosticator_status, prop_year_status) %>%
  ggplot(aes(x = year, y = prop_year_status)) +
    geom_point() +
    geom_smooth() +
    facet_wrap(~ prognosticator_status) +
    scale_y_continuous(labels = scales::percent_format(),
                       limits = c(0, 1)) +
    labs(x=NULL, y=NULL,
         title = str_wrap("Inanimate creatures are becoming better prognosticators over time", 60),
         subtitle = "Comparing yearly prediction accuracy among real and fake prognosticators",
         caption = str_wrap("Source: Countdown to Groundhog Day and U.S. National Ocean and Atmospheric Administration. Limited to years with at least 5 predictions per group.", 100)) +
    theme_bw() + 
    theme(plot.title.position = "plot",
          plot.title = element_text(face="bold"))
```
