---
title: "English Education"
format: html
---


# Exploration
 
*Use what you know so far to explore the dataset. If you are just starting out, write about what you want to do with the data. Try to stretch a bit every week.*

```{r}
# load packages
  library(tidyverse)
  library(tidytuesdayR)   # install.packages("tidytuesdayR") to get new data

  english_education <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-01-23/english_education.csv')

```

```{r}
  english_education %>%
    ggplot(aes(x=education_score)) +
    geom_histogram()
```







# Review

## Find and run

*Search the Twitter hashag [#TidyTuesday](https://twitter.com/hashtag/TidyTuesday?src=hashtag_click) for an entry that catches your eye (and includes a link to their code). Include links to the details and run their code.*

| Name | Details |
|:--   |:--      |
| Author | Replace |
| Tweet | [url](https://twitter.com/...) |
| Code | [url](https://github.com/...) |

```{r entry}
# insert their code

```

## Notes

*What did you learn from their approach? What would you do differently?*

