---
title: "The Office 1"
author: "Your name"
output: github_document
---

```{r load-packages, message = FALSE}
# credit to https://datasciencebox.org/index.html
library(tidyverse)
library(tidymodels)
library(schrute)
library(lubridate)
```

In this two-part exercise we will use `theoffice` data from the [**schrute**](https://bradlindblad.github.io/schrute/) package to predict IMDB scores for episodes of The Office. In the first part, we'll work together to prepare the data, specify a model, and create a recipe. In the second part, you will fit and evaluate the model on your own for homework.

```{r}
glimpse(theoffice)
```

Fix `air_date` for later use.

```{r}
theoffice <- theoffice %>%
  mutate(air_date = ymd(as.character(air_date)))
```

We will

-   engineer features based on episode scripts
-   train a model
-   perform cross validation
-   make predictions

```{r}
theoffice %>%
  distinct(season, episode)
```

### Exercise 1 - Calculate the percentage of lines spoken by Jim, Pam, Michael, and Dwight for each episode of The Office.

```{r lines}
office_lines <- theoffice %>%
  group_by(season, episode) %>%
  mutate(
    n_lines = n(),
    lines_jim = sum(character=="Jim") / n_lines,
    lines_pam = sum(character=="Pam") / n_lines,
    lines_dwight = sum(character=="Dwight") / n_lines,
    lines_michael = sum(character=="Michael") / n_lines
  ) %>%
  ungroup() %>%
  select(season, episode, episode_name, contains("lines")) %>%
  distinct(season, episode, episode_name, .keep_all = TRUE)
```

### Exercise 2 - Identify episodes that touch on Halloween, Valentine's Day, and Christmas.

```{r special-episodes}
theoffice <- theoffice %>%
  mutate(text = tolower(text))

#theoffice$text <- tolower(theoffice$text)

halloween_episodes <- theoffice %>%
  filter(str_detect(text, "halloween")) %>%
  count(episode_name) %>%
  filter(n > 1) %>%
  mutate(halloween = 1) %>%
  select(-n)

valentine_episodes <- theoffice %>%
  filter(str_detect(text, "valentine")) %>%
  count(episode_name) %>%
  filter(n > 1) %>%
  mutate(valentine = 1) %>%
  select(-n)

christmas_episodes <- theoffice %>%
  filter(str_detect(text, "christmas")) %>%
  count(episode_name) %>%
  filter(n > 1) %>%
  mutate(christmas = 1) %>%
  select(-n)
```

### Exercise 3 - Put together a modeling dataset that includes features you've engineered. Also add an indicator variable called `michael` which takes the value `1` if Michael Scott (Steve Carrell) was there, and `0` if not. Note: Michael Scott (Steve Carrell) left the show at the end of Season 7.

```{r office-df}
office_df <- theoffice %>%
  select(season, episode, episode_name, imdb_rating, director, writer, 
         total_votes, air_date) %>%
  distinct(season, episode, .keep_all = TRUE) %>%
  left_join(halloween_episodes, by = "episode_name") %>%
  left_join(valentine_episodes, by = "episode_name") %>%
  left_join(christmas_episodes, by = "episode_name") %>%
  replace_na(list(halloween = 0, valentine = 0, christmas = 0)) %>%
  mutate(michael = case_when(
      season > 7 ~ 0,
      TRUE ~ 1
    )) %>%
  mutate(across(halloween:michael, as.factor)) %>%
  left_join(office_lines, by=c("season", "episode", "episode_name"))
```

### Exercise 4 - Split the data into training (75%) and testing (25%).

```{r split}

```

### Exercise 5 - Specify a linear regression model.

```{r model}

```

### Exercise 6 - Create a recipe that updates the role of `episode_name` to not be a predictor, removes `air_date` as a predictor, uses `season` as a factor, and removes all zero variance predictors.

```{r recipe}

```

### Exercise 7 - Build a workflow for fitting the model specified earlier and using the recipe you developed to preprocess the data.

```{r workflow}

```
